\section{Thesis Contributions}

This work has three main goals: research, design, and implementation.

\subsection{Research}

The research goal is to conduct an analysis of compute resource usage at our department. This includes describing different use-cases users have for the system in place and creating typical usage scenarios for both users and administrators. User interviews were conducted with students experienced with the current solution, covering their workflows, data, experiment outputs, hardware and software needs, limiting factors, and pain points.

\todo{Forward-reference the chapter/section where interview findings are presented in detail.}

\subsection{Design}

The design goal is to design solutions that support the identified use-cases. The solution should:
\begin{itemize}
    \item be able to exist within the existing infrastructure,
    \item be easily scalable,
    \item provide additional services,
    \item support modern AI/ML practices.
\end{itemize}

\subsection{Implementation}

After providing solutions that address users' needs, the goal is to select a solution and implement it. \todo{Briefly state or forward-reference the selection criteria used to choose the final design (e.g., compatibility with existing infrastructure, scalability, operational overhead).} Specifically, this thesis contributes:

\begin{itemize}
    \item A modular, Kubernetes-native platform for AI research.
    \item An intelligent scheduling layer for resource-efficient batch processing.
    \item A distributed executor for scaling research workloads.
\end{itemize}

\todo{Consider adding a brief closing paragraph that maps each contribution to the thesis structure (Chapter X covers research, Chapter Y covers design, etc.).}
