\section{Problem statement}

The complexity of managing distributed AI workloads and the difficulty of providing a ``single-pane-of-glass'' interface for researchers without heavy DevOps overhead.

\subsection{Data and storage management}

Machine learning often requires large datasets that need to be present for training and evaluating models. Users transfer their datasets onto the machines manually ahead of computation. The datasets then reside on the machines indefinitely. Two problems stem from this:

\begin{itemize}
    \item When the machine a user has all their data on is being used by someone else, the user needs to either wait for the other job to finish or move the data elsewhere, causing further strain on the machine.
    \item Users are often not mindful about deleting their unused data. This causes the file system to fill up and requires periodical clean-ups by the administrator.
\end{itemize}

\subsection{Environment management}

Users manage their environments, packages, and libraries through Conda. As with data, however, these environments are not shared across the machines. This means that whenever users want to switch machines they need to take their Conda environments with them. The result is that the same packages are installed multiple times across the machines, increasing file system utilization further.

\subsection{Scheduling}

Whenever users want to use the machines for compute they are supposed to post into a Teams channel. The post should contain information about their job --- mainly the length of the computation and the machine(s) they are selecting. Many times, however, the time estimate is inaccurate, or even worse --- users do not post about their job in the channel and just start using the machines. This results in users needing to check each machine individually to confirm whether it is available. There is no centralized monitoring solution, so users manually connect to each node and check its utilization.
