\section{Batch vs. Interactive Computing}
\label{sec:batch-vs-interactive}

The challenge of scheduling high-throughput AI jobs alongside interactive researcher notebooks.

\todo{Expand this into a full introductory paragraph explaining the tension between batch and interactive workloads and why it matters for the platform design.}

\subsection{Common use-cases}

Through user interviews with students experienced with the current solution, several common use cases were identified:

\todo{Reference the methodology section or appendix where interview details are described (sample size, interview protocol, etc.).}

\begin{enumerate}
    \item \textbf{Script-based batch execution.} Accessing the servers via SSH, copying scripts and datasets onto the machine(s), and running the scripts directly.
    \item \textbf{Third-party experiment management.} Accessing the servers via SSH, running third-party daemons on the servers, and managing experiments from a web interface. This allows users to run distributed scripts, optimize hyper-parameters of the models, and provides useful metrics about the training process.

\todo{Name the specific third-party tools mentioned in interviews (e.g., Weights \& Biases, MLflow, Optuna) to make the use-case concrete.}
    \item \textbf{Remote development environment.} Using the servers as remote development environments. Users SSH onto the servers and work on their research remotely. This centralization is important for them as they can use thin clients, and if they need to run any quick GPU-heavy calculations to validate their work, there is no need to move to another machine.
\end{enumerate}

\todo{Add subsections expanding on the scheduling challenges for each use-case type and how they conflict (e.g., a long-running notebook holding a GPU vs.\ a batch training job queuing for it). Forward-reference Volcano's queue/priority mechanisms as the proposed solution.}
